---
title: "Data import"
output: github_document
---


This file is for doing data import.

##Import data sets

```{r}
library(tidyverse)
library(readxl)
library(haven)
```

Import our first data set
(we went back and amended this data import to include na setting to )
```{r}
litters_df = 
  read_csv(file = "data/FAS_litters.csv",
           na = c(".", "NA", "")
           )

litters_df = janitor::clean_names(litters_df)
```

Import your second data set
Janitor cleans data names for sameness (no capitals, same formatting, etc.)

```{r}
pups_df=
  read_csv(file="data/FAS_pups.csv", 
           skip = 3,
           na = c(".", "NA", "")
           )

pups_df = janitor::clean_names(pups_df)
```

## Look at the data
```{r}
litters_df
pups_df
```

Head or Tails

```{r}
head(litters_df, 29)

tail(pups_df, 5)
```

skimming? gives very quick complete breakdown of whats in the dataset (quick summary). Includes mean, missing, sd, percentiles, variable names etc.

```{r}
skimr::skim(pups_df)
```




# Oh excel...

Jenny Bryan made 'readxl' to solve our problems.

```{r}
mlb_df = read_excel("data/mlb11.xlsx")
```

Look at the data!

```{r}
mlb_df
```


Import FOTR words.

```{r}
fotr_df = 
  read_excel("data/LotR_Words.xlsx",
             range = "B3:D6")

fotr_df
```

what about two towers? (Same file different table within same excel file.)

```{r}
tt_df = 
  read_excel("data/LotR_Words.xlsx",
             range = "F3:H6")

tt_df
```

## Import SAS

Read in the PULSE set

```{r}
pulse_df = 
  read_sas("data/public_pulse_data.sas7bdat")

pulse_df = janitor::clean_names(pulse_df)
```

Import pups with 'read.csv()'

```{r}
pups_df_base = 
  read.csv("data/FAS_pups.csv")

pups_df_base
```

problem with read.csv --> won't show: lazy about variable names; super slow for big dataset.