---
title: "02_data_manipulation"
output: github_document
---

This file is for doing data manipulation.

##Import data sets

```{r}
library(tidyverse)
library(readxl)
library(haven)
```

Import Litters
```{r}
litters_df = 
  read_csv(file = "data/FAS_litters.csv",
           na = c(".", "NA", "")
           )

litters_df = janitor::clean_names(litters_df)
```

Import pups

```{r}
pups_df=
  read_csv(file="data/FAS_pups.csv", 
           skip = 3,
           na = c(".", "NA", "")
           )

pups_df = janitor::clean_names(pups_df)
```

## `Select` some variables

```{r}
select(litters_df, group, litter_number)

select(litters_df, group:gd_of_birth)

select(litters_df, starts_with("gd"))

select(litters_df, starts_with("gd"), group)

select(litters_df, GROUP = group, everything()) #renaming the variable thats on the right of = what is on the left. 

select(litters_df, -starts_with("gd"))
```

a couple of alternatives

```{r}
rename(litters_df, GROUP = group)

#changing position/order of columns
relocate(litters_df, litter_number, gd0_weight)
```

In the oups data, select litter number, sex and PD ears.

```{r}
select(pups_df, litter_number, sex, pd_ears)
```

## `filter` some rows

```{r}
filter(litters_df, gd_of_birth == 20)

filter(litters_df, pups_born_alive <= 5)

filter(litters_df, pups_born_alive != 6)

filter(litters_df, group %in% c("Con7", "Con8"))

filter(litters_df, group %in% c("Con7", "Con8"), pups_born_alive < 7)

#OR function
filter(litters_df, pups_born_alive < 4 | pups_born_alive < 7)

#getting rid of all missing data/NA data in specific columns
drop_na(litters_df, gd0_weight)

```

In the pups dataset, keep only rows where PD walk is less than 11 and sex is 2

```{r}
filter(pups_df, pd_walk < 11, sex == 2)
```


## `mutate` to change some columns

let's add and change some columns

```{r}
mutate(
  litters_df, 
  wt_gain = gd18_weight - gd0_weight, #named variable and variable definition
  wt_gain_squared = wt_gain^2, #You can modify variables immediatly after creating them
  group = str_to_lower(group)) #make everything in this variable lowercase
```

In the pups dataset, create a variable that subtracts 7 from PD walk

```{r}
mutate(
  pups_df,
  pd_walk_minus_7 = pd_walk - 7
)
```


## `arrange` the data

```{r}
arrange(litters_df, pups_born_alive)
arrange(litters_df, desc(pups_born_alive))
```

##Do multiple steps

This is bad:

```{r}
litters_df = read_csv("data/FAS_litters.csv", na=c("", "NA", "."))
litters_rename_df = janitor::clean_names(litters_df)
litters_with_gd_df = select(litters_rename_df, group, starts_with("gd"))
litters_no_na_df = drop_na(litters_with_gd_df)
litters_wt_gain_df = mutate(litters_no_na_df, wt_gain = gd18_weight - gd0_weight)
```

THIS is good: pipe operator takes whatever is in the last line and tells the next operator to amend that object. Basically "and then".

```{r}
litters_df=
  read_csv("data/FAS_litters.csv", na=c("", "NA", ".")) |> 
  janitor::clean_names() |> 
  select(group, starts_with("gd")) |> 
  drop_na() |> 
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)
  )
```


load pups, clean names, drop missing, keep litter_number and pd variables, add pd walk -7

```{r}
pups_df=
  read_csv("data/FAS_pups.csv", skip = 3, na = c("", "NA",".")) |> 
  janitor::clean_names() |> 
  drop_na() |>
  select(litter_number, starts_with("pd")) |> 
  mutate(
    pd_walk_minus_7 = pd_walk - 7
  )
```


sometimes the first thing isn't a data frame

```{r}
litters_df |> 
  filter(group %in% c("con7", "con8")) |> 
  lm(gd18_weight ~ gd0_weight, data = _)
```

`select` vs `pull`

```{r}
litters_df |> 
  select(group)

litters_df |> 
  pull(group)
```

never do `$`, amend using mutate, select, pull, etc. 

```{r}
litters_df$group
```

